---
title: "Homework 3"
author: Tehya Boggs
date: May 28, 2025
format: html
editor: visual
---

### GitHub Repository Link: <https://github.com/tboggs1219/ENVS-193DS_homework-03.git>

```{r setup, message = FALSE, warning = FALSE}

# install.packages("here")
# install.packages("readxl")

# read in packages
library(tidyverse)
library(here)
library(flextable)
library(janitor)
library(readxl)
library(dplyr)

# read in data
data <- read_excel(
  # data path
   here("data", "personal-data.xlsx"))
```

### Problem 1. Personal data

#### a. Data summarizing

My data can be summarized by calculating mean energy level to compare between days, with each day having variable caffeine intake, sleep quality, sleep duration, etc.: mean daily energy level will be my response variable, with potential explanatory variables including those aforementioned (caffeine intake, sleep quality, etc.). Calculating the average for each day will allow for comparison given the fact that there are daily differences between explanatory variables.

#### b. Visualization

```{r visualization}

data$`date (yyyy-mm-dd)` <- as.Date(data$`date (yyyy-mm-dd)`) # classify date column as date

data_clean <- data %>% # using data df
  group_by(`date (yyyy-mm-dd)`) %>% # group rows by date
  summarise(
    # new column calculating mean energy per day
    avg_energy = round(mean(`energy level (1-10)`), 1),
    # new column showing total caffeine     intake per day
    total_caffeine = max(`total caffeine intake (mg)`),
    # new column showing single sleep quality rating for that day
    sleep_quality = max(`sleep quality (1-10)`)
  ) %>% 
  # clean column names
  clean_names() 

# plot using summarized data
ggplot(data = data_clean,
       aes(x = factor(sleep_quality), # sleep quality score on the x-axis
           y = avg_energy)) + # avg energy on the y-axis
 
   # first layer: boxplot
  geom_boxplot(fill = "lightblue3", color = "darkblue") +
  
  # second layer: jitter showing underlying data, each data point represents one day
  geom_jitter(width = 0.3, # set horizontal jitter
              height = 0, # set vertical jitter
              shape = 21, # change shape to support fill/color arguments
              fill = NA, # no fill
              color = "royalblue4", # change outline
              stroke = 1.1) + # outline thickness
 
  labs(
    # chart title
    title = "Average Energy Level by Sleep Quality",
    # x-axis label
    x = "Sleep Quality (1-10)",
    # y-axis label
    y = "Average Energy Level (1-10)"
  ) +
    theme_minimal() # change theme from default


```

#### c. Caption

The above figure depicts average energy level by sleep quality, with the box plot layer showing median energy and IQR across sleep quality ratings. The underlying data, represented by open points, show daily averages, with each point symbolizing one day of observations. Energy level was recorded three times per day (in the morning, afternoon, and evening) in order to calculate daily averages. Generally, as sleep quality increases, average energy level also increase, evidenced by the increasing median energy levels across sleep quality ratings.

#### d. Table presentation

```{r table presentation}

# display summarized data table using data_clean
flextable(data_clean) %>% 
  # change column names
  set_header_labels( 
    date_yyyy_mm_dd = "Date",
    avg_energy = "Mean Daily Energy Level (1-10)",
    total_caffeine = "Total Caffeine Intake (mg)",
    sleep_quality = "Sleep Quality (1-10)"
  ) %>% 
  autofit()

```

### Problem 2. Affective visualization

#### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).

Affective visualization methods seek to implement creative strategies in order to have an emotional impact on viewers. In the context of my data, featuring daily energy level according to caffeine intake and sleep quality, affective visualization could take the form of a calendar heat map, similar to that of temperature heat maps that show temperature averages by color. Instead, cell color would represent average energy level, with each cell representing one day. Icons depicting caffeine intake and sleep quality would be overlayed over each cell, potentially with a coffee cup indicating how many energy drinks were consumed per day, and icons to indicate sleep quality categorized into high (7-10), moderate (4-6), and low (1-3) designations.

#### b. Create a sketch (on paper) of your idea.

#### c. Make a draft of your visualization.

#### d. Write an artist statement.

The above heat map illustrates daily average energy level on a monochromatic color scale, with darker colors representative of higher energy on a scale from one to ten, also featuring mugs and stars to delineate caffeine intake and sleep quality. The figure draws inspiration from temperature heat maps, such as Lorraine Woodruff-Long's warming strips quilt, and was created using acrylic paint on canvas.

### Problem 3. Statistical critique

#### a. Revisit and summarize

"Statistical analysis of the patterns and characteristics of connected and autonomous vehicle involved crashes" by Xu et al. investigates the severity of connected and autonomous vehicle (CAV) involved traffic collisions dependent on multiple variables, ranging from CAV driving mode, accident location, the presence of roadside parking, and more. The authors developed a binary logistic regression model to describe how such independent factors influence the severity of CAV involved traffic collisions, with severity determined on a binary scale involving either property-damage-only (PDO) crashes or crashes with reported injury or pain. These independent factors were coded on a binary scale with a value of 0 or 1 to indicate a variables involvement in a collision. The output of this statistical test provides a coefficient that describes the strength of the association of the independent variable to the response variable, as well as a p-value indicating the statistical significance of such correlation. Authors presented test results for the crash severity model in a table.

#### b. Visual clarity

The data effectively represents the binary logistic regression model for CAV collision severity given that it only includes predictor variables that are statistically significant, despite other factors having been investigated in the paper's analysis.

#### c. Aesthetic clarity

Authors combated visual clutter by bolding column titles, and presenting coefficient values in the first column after variable name for greater readability. Summary statistics indicating log transformed odds and odds ratio values are also somewhat separate from variable statistics, increasing the clarity of the test output.

#### d. Recommendations

It is recommended that the variable names be simplified for viewer to more easily interpret their meaning on a binary scale. For example, the variable "CAV stopped or not" could be changes to "CAV stopped" to facilitate understanding that a coded value of 1 indicates the affirmation of that variable, as in the CAV is stopped. Further more, "Mode" could be "Autonomous Driving" so that driving mode can be inferred without Table 1 in the article, which serves as a key for variable categories and meanings.
